---
title: "C3_WK4_A1"
date: "September 3, 2016"
output: html_document
---
# Assignment
The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set. 

The goal is to prepare tidy data that can be used for later analysis. 

You will be graded by your peers on a series of yes/no questions related to the project. You will be required to submit: 

- 1) a tidy data set as described below, 
- 2) a link to a Github repository with your script for performing the analysis,
- 3) a code book that describes the variables, the data, and any transformations or work that you performed to clean up the data called CodeBook.md. 

# Load Data & Explore

```{r}
library(plyr)
library(data.table)

setwd("~/Documents/R Working Directory/Course_3_Getting_and_Cleaning_Data") ## Set the Working Directory

if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip" ### Provide the URL of where to get the file from 

download.file(fileUrl,destfile="./data/Dataset.zip",method="curl") ### Download the file and place it in a directory for analysis
unzip(zipfile="./data/Dataset.zip",exdir="./data") ### Unzip the ZIP file with the built in module
```

We need to get the SUBJECT Training & Test files to complete our activities.  Based on the information provided we will need to get the TRAIN & TEST files for SUBJECT, X & Y.

```{r}
setwd("~/Documents/R Working Directory/Course_3_Getting_and_Cleaning_Data/data/UCI HAR Dataset")
subjectTrain = read.table('./train/subject_train.txt',header=FALSE)
xTrain = read.table('./train/x_train.txt',header=FALSE)
yTrain = read.table('./train/y_train.txt',header=FALSE)

subjectTest = read.table('./test/subject_test.txt',header=FALSE)
xTest = read.table('./test/x_test.txt',header=FALSE)
yTest = read.table('./test/y_test.txt',header=FALSE)

Activity  = read.table('./activity_labels.txt',header=FALSE) ## Load the Activity Labels
Features = read.table('./features.txt',header=FALSE) ## Load the Feature Description 
```

# Clean & Transform Data
Now that the data has been tagged, I will proceed to join the data into a single table and label the columns as required
```{r echo = FALSE}

xDataSet <- rbind(xTrain, xTest)
yDataSet <- rbind(yTrain, yTest)
subjectDataSet <- rbind(subjectTrain, subjectTest)
```

##Extract only the measurements on the mean and standard deviation for each measurement
```{r}
xDataSet_mean_std <- xDataSet[, grep("-(mean|std)\\(\\)", Features[, 2])]
names(xDataSet_mean_std) <- Features[grep("-(mean|std)\\(\\)", Features[, 2]), 2] 
```

##Apply descriptive activity names to name the activities in the data set
```{r}
yDataSet[, 1] <- Activity[yDataSet[, 1], 2]
names(yDataSet) <- c("Activity")
yDataSet[1:5,]
names(subjectDataSet) <- c("Subject")
```
## Create a single data set
```{r}
AggregatedDataSet <- cbind(xDataSet_mean_std, yDataSet, subjectDataSet)
```

## Label the Columns with more descriptive names based on Features.Txt
```{r}
names(AggregatedDataSet) <-gsub('tBodyAcc-',"Time_Domain_Body_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tGravityAcc-',"Time_Domain_Gravity_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyAccJerk-',"Time_Domain_Linear_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyGyro-',"Time_Domain_Angulaur_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyGyroJerk-',"Time_Domain_Angulaur_Acceleration_Velocity_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyAccMag-',"Time_Domain_Body_Acceleration_Magintude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tGravityAccMag-',"Time_Domain_Gravity_Acceleration_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyAccJerkMag-',"Time_Domain_Linear_Acceleration_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyGyroMag-',"Time_Domain_Angulaur_Acceleration_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('tBodyGyroJerkMag-',"Time_Domain_Angulaur_Acceleration_Velocity_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyGyroJerkMag-',"Frequency_Domain_Angulaur_Acceleration_Velocity_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyAccJerkMag-',"Frequency_Domain_Linear_Acceleration_Magnitude_",names(AggregatedDataSet))                             
names(AggregatedDataSet) <-gsub('fBodyGyroMag-',"Frequency_Domain_Angulaur_Acceleration_Magnitude_",names(AggregatedDataSet))                                       
names(AggregatedDataSet) <-gsub('fBodyGyroJerkMag-',"Frequency_Domain_Angulaur_Acceleration_Velocity_Magnitude_",names(AggregatedDataSet))    
names(AggregatedDataSet) <-gsub('fBodyAccJerk-',"Frequency_Domain_Linear_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyAccJerkMag-',"Frequency_Domain_Linear_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyGyroMag-',"Frequency_Domain_Angulaur_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyAccMag-',"Frequency_Domain_Body_Acceleration_Magintude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyGyro-',"Frequency_Domain_Angulaur_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyAcc-',"Frequency_Domain_Body_Acceleration_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyBodyAccJerk-',"Frequency_Domain_Linear_Acceleration_",names(AggregatedDataSet))                            
names(AggregatedDataSet) <-gsub('fBodyBodyGyroMag-', "Frequency_Domain_Angulaur_Acceleration_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('fBodyBodyGyroJerkMag-',"Frequency_Domain_Angulaur_Acceleration_Velocity_Magnitude_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('mean',"Mean_",names(AggregatedDataSet))
names(AggregatedDataSet) <-gsub('std',"Standard_Deviation_",names(AggregatedDataSet))
```

## Create a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r}
FinalDataSet<-aggregate(. ~Subject + Activity, AggregatedDataSet, mean)  ## Create a new data set that finds the average for each of the different Subjects and Activities
write.table(FinalDataSet, file = "./tidydata.txt",row.name=FALSE)
```
